---
title: "MSNA Dashboard Template"
output: github_document   
---


<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![contributions welcome](https://img.shields.io/badge/contributions-welcome-brightgreen.svg?style=flat)](https://github.com/dwyl/esta/issues)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Generic badge](https://img.shields.io/badge/STATUS-UNDER%20DEVELOPMENT-%23007CE0)](https://shields.io/)


<!-- badges: end -->

The goal of MSNAdashboardTemplate is to ...

<!-- TOC -->

## Installation

You can install the development version of MSNAdashboardTemplate like so:

``` {r eval =FALSE}
devtools:install_github("MSNAdashboardTemplate")
```

## Step by step guideline
#### STEP000:: Call the library

```{r eval =FALSE}
library(MSNAdashboardTemplate)
# library(analysistools)
```

#### STEP 001. Preapare the kobo tool
At the very beginning where you are developing the KOBO tool please make sure you have used  OCHA COD name in your KOBO tool’s name column. You’re free to modify the label as desired, but please avoid changing the name column. You can download OCHA COD using the `download_hdx_adm()` function, which automatically retrieves the data from HDX. To obtain a name that can be directly copied to your KOBO tool, use `snakecase::to_snake_case()`. Here’s an example:

```{r eval =FALSE}
admin_boundary <- MSNAdashboardTemplate::download_hdx_adm(country_code = "lbn",admin_level = 2,df_index = 2)  ## Downloading OCHA-COD admin 2 data for Lebanon
admin_boundary <- admin_boundary |> mutate(
  name = snakecase::to_snake_case(admin2Name), ## Change admin2Name as necessary for your country
  `label::fr` = admin2Name
) |> as.data.frame()|> select( `label::en`,name)

```
The code above should generates a table like following, that can be copied directly into the KOBO choices tab. While you’re free to modify the label::en, please refrain from changing the name, as it will be used for automation/upcoming functions that will be available soon.

#### STEP 002. Create DAP for the dashboard

```{r eval =FALSE}

survey<- readxl::read_excel("data-raw/kobo_tool.xlsx",sheet = "survey")

HH_data<- readxl::read_excel("data-raw/REACH_LBN_cleaned_data.xlsx",sheet = "HH_data")
INDV_data<- readxl::read_excel("data-raw/REACH_LBN_cleaned_data.xlsx",sheet = "INDV_data")

list_of_dataset <- list(
  HH_data = HH_data,
  INDV_data = INDV_data
)


MSNAdashboardTemplate::write_dap_for_dashboard(list_of_dataset = list_of_dataset,
                        kobo_survey = survey,
                        column_name_for_populaion_group = "pop_group")

```
Once you run the above code, you will have a excel file with two tab. One `dap_for_dashboard` which is looks like - 

```{r eval =T,warning=F ,echo=F}
 readxl::read_excel("data-raw/data_diagnosis.xlsx",sheet = 1)
```

Here you should fill out the `sector` from dropdown and `indicator` name aganist main_variable.  Please note that if you keep `indicator` empty then it wont be showing in the dashboard. 
In the second tab, please select the `strata` and `admin_level` from dropdown menue 


```{r eval =T,warning=F ,echo=F}
 readxl::read_excel("data-raw/data_diagnosis.xlsx",sheet = 2)
```

#### STEP 003. Create list of mismatched strata
To join the map data with MSNA data, the strata name from the dataset should be same as admin boundary name in the OCHA boundary. You can use `write_mismatched_strata()` to compare the the admin names. 


```{r eval =FALSE}
cntry_adm_3 <- MSNAdashboardTemplate::download_hdx_adm(country_code = "lbn",admin_level = 3,df_index = 2,keep_zip = T)

strata_checking <- readxl::read_excel("data-raw/data_diagnosis.xlsx",sheet = "strata_checking")
df <- readxl::read_excel("data-raw/REACH_LBN_cleaned_data.xlsx",sheet = "HH_data")


MSNAdashboardTemplate::write_mismatched_strata(ocha_cod_level_3 = cntry_adm_3,
                        cod_admin1_name = "admin1Name",
                        cod_admin2_name = "admin2Name",
                        cod_admin3_name = "admin3Name",
                        strata_checking_sheet = strata_checking,
                        hh_data=df,
                        population_group_column_name_strata_checking = "pop_group" )



```

The function will give an excel file which will look like the following table. Please make sure you have filled the `ocha_cod_name` from the drop down menu. 

```{r eval =T,warning=F ,echo=F}
 readxl::read_excel("data-raw/admin_mismatch_fix.xlsx",sheet = 1)
```


#### STEP 004. Rename OCHA name 
Once you fill the output file from `write_mismatched_strata` function then we will need to apply the changes to either dataset or to OCHA cod.However as MSNA data mostly contain multiple loop so its better to rename the ocha boundaries. You can apply `write_renamed_ocha_admin_name` so write the renamed OCHA boundary as shapefile to use in the dashboard. 
```{r eval =FALSE}
MSNAdashboardTemplate::write_renamed_ocha_admin_name(ocha_cod_level_3 = cntry_adm_3,
                              cod_admin1_name = "admin1Name",
                              cod_admin2_name = "admin2Name",
                              cod_admin3_name = "admin3Name",
                              strata_checking_sheet =    strata_checking,
                              mistached_list = mistached_list,
                              output_path = "data-raw/OCHA_renamed_adm3.shp")
```





#### STEP 005. Recreate_analysis_file
You can do analysis using any tools or package or whatever you want but make sure you have `key_index` and `state`

```{r eval =FALSE}

```

#### step 005. Data preparation for dashboard
[to be done -
1. Identify grouping variables from key index
2. Identify main variable from key index
3. Identify analysis type from key index
4. Identify Question type from key index/kobo/dap]
5. Joining DAP information




#### STEP 006. Create text file

The function `create_text_input()` will create the an excel file with two tab (1. `README`, 2. `text_input`). Please read the readme and then fill the `text_input` tab. 

```{r eval =FALSE}
create_text_input(output_path = "[PATH]")
```


#### STEP 005. Create summary
Ideally it will check all the module for the dashboard It should provide list of error/issue to fix before running the create_dashboard 

```{r eval =FALSE}

```



#### STEP 007. Create dashboard

```{r eval =FALSE}

MSNAdashboardTemplate::create_dashboard(country = "Iraq",
                               assessment_name  = "McNA",
                               year = 2024,
                               analysis_file = analysis_file,
                               key_index_column= key_index_column, 
                               stat_column = stat_column
                               ) 
```



### How to customize the dashboard? 

In case you want to customize the dashboard, you can follow the following steps -

#### Step 0.1.1 Clone the repo

To customized the dashboard, you first need to clone the repository from the [github](https://github.com/mhkhan27/MSNAdashboardTemplate). 


#### Step 0.1.1 Understanding the repo
Once you are done with the repo, now you need to understand the structure of repo. The repository is created using `golem` package. 

##### What is golem?

The `golem` package is a framework for building and deploying production-ready Shiny applications in R. Shiny is a web application framework for R that allows users to create interactive web applications using R code. However, while Shiny is great for creating interactive prototypes, it is not always well-suited for building production-ready applications with robust performance, security, and scalability.


This is where`golem` comes in. `Golem` provides a framework for organizing and structuring your Shiny code to create scalable and maintainable applications. It includes a set of best practices and conventions for building Shiny applications, such as separating the UI (user interface) and server logic into separate files, using reactive programming to minimize data processing, and leveraging package management to simplify dependency management.


In addition to these best practices, `golem` also provides a set of tools for testing, debugging, and deploying Shiny applications. For example, it includes a command-line interface for creating and managing application templates, as well as tools for managing application configuration, logging, and error handling.

Overall, the `golem` package is a powerful tool for building robust, scalable, and maintainable Shiny applications in R.


##### Repository Stucture 

The Golem package in R is designed to facilitate the creation of production-ready Shiny applications. It is organized into several subdirectories, each of which serves a specific purpose:

1. `R/:` This directory contains the R code for your application. You should place your application logic, including any functions you write, in this directory.

2. `data/:` This directory is used to store any data files that your application needs to function. You can also store any other resources that your application requires in this directory.

3. `www/:` This directory is used to store any static files that your application requires, such as images, stylesheets, or JavaScript files.

4. `inst/:` This directory is used to store any additional files that your application needs to function, such as configuration files or documentation.

5. `tests/:` This directory is used to store any test files for your application. You can write unit tests for your application logic in this directory.

6. `man/:` This directory contains the documentation for your application. You should document your functions and other objects in this directory using the Roxygen2 syntax.

7. `NAMESPACE:` This file specifies the package's exported functions and other objects.

8. `DESCRIPTION:` This file contains metadata about your package, including its name, version, and dependencies.

9. `README.md:` This file contains information about your package, including how to install and use it

```{r eval =T, echo=FALSE}
 fs::dir_tree()
```

##### Understanding Module 

With in the `R/` folder you will scripts started with `mod_` which I will be calling module. Each module represent a `tab` in the dashboard. Within the scripts, each module have two parts, 1. user interface (ui) 2. server. 

#### Step 0.1.1 Removing any module
If you want to remove any module/tab from the dashboard, you can easily do it from `app_ui.R` and `app_server.R`. All you have to do is delete the module from both `app_ui` and `app_server`


#### Step 0.1.1 Adding any module

If you want to add a new tab then you will need to create a module first and then add the module to the `app_ui.R` and `app_server.R`

```{r eval =F}
## to create the module
golem::add_module(name = "name_of_module1", with_test = TRUE) # Name of the module

```

#### Step 0.1.1 Changing any module
You can change a specific tab without creating error in out tab by editing existing module. just make the changes in `mod_name_of_module1` and then you are all set. [Make sure you have made changes in both ui and server]


#### Step 0.1.1 Check the tests

If you made any changes in the dashboard, please make sure you have run the tests before deploying the app. It will make sure the the changes didn't change anything in other tab or in the app. You can hit `Ctrl + Shift + T` to run the tests   
 


#### Step 0.1.1 CMD check 
It is also important to run the CDM check. If your CDM check fails then the app may runs in local but in the shinyserver it might not run. You can hit  `Ctrl + Shift + E` to run the CMD checks 


#### Step 0.1.1 Run the app

```{r eval =FALSE}


MSNAdashboardTemplate::create_dashboard(country = "Iraq",
                               assessment_name  = "McNA",
                               year = 2024,
                               analysis_file = analysis_file,
                               key_index_column= key_index_column, 
                               stat_column = stat_column
                               ) 
```

